---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# IDConverter

<!-- badges: start -->
[![Lifecycle: stable](https://img.shields.io/badge/lifecycle-stable-brightgreen.svg)](https://lifecycle.r-lib.org/articles/stages.html)
[![](https://cranlogs.r-pkg.org/badges/grand-total/IDConverter?color=orange)](https://cran.r-project.org/package=IDConverter)
[![R-CMD-check](https://github.com/WangLabCSU/IDConverter/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/WangLabCSU/IDConverter/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The goal of IDConverter is to convert identifiers between biological databases.
Currently, I mainly use it for promoting cancer study.

## Installation

Install the latest version of IDConverter in GitHub with:

``` r
remotes::install_github("WangLabCSU/IDConverter")
```

Or Gitee (better in China, but not updated anymore):

```r
remotes::install_git("https://gitee.com/WangLabCSU/IDConverter")
```

## Available features

**ID conversions**:

- `convert_custom()` - Convert custom database identifiers.
- `convert_icgc()` - Convert ICGC identifiers.
- `convert_pcawg()` - Convert PCAWG identifiers.
- `convert_tcga()` - Convert TCGA identifiers.
- `convert_hm_genes()` - Convert human/mouse/worm gene IDs between Ensembl and Hugo Symbol system.
- `convert_hm_orthologs()` - Convert gene IDs between human and mouse via Ensembl orthology (e.g., TP53 ↔ Trp53).

**Annotation tables**:

- `ls_annotables()` - List available annotation tables from [annotables](https://github.com/stephenturner/annotables).
- `load_data()` - Download and load annotation data from Zenodo (fixed version) or from local cache.
- `build_annotables()` - Build up-to-date gene annotation tables directly from Ensembl BioMart
  (supports 11 organisms, requires `biomaRt`).
- `resolve_gene_aliases()` - Resolve outdated/alternative gene symbols (aliases) to official
  symbols via Ensembl synonym data.

**Utilities**:

- `pair_gdc_samples()` - Pair tumor-normal samples from GDC manifest files.
- `parse_gdc_file_uuid()` - Parse Metadata from GDC Portal File UUID.
- `filter_tcga_barcodes()` - Filter TCGA Replicate Sample Barcodes.

## Examples

```{r}
library(IDConverter)
```

To follow the CRAN policy, I have to set `tempdir()` as default data path,
however, I recommend you set the data path to a specified path with
`options(IDConverter.datapath)`.

e.g.,

```{r}
options(IDConverter.datapath = system.file("extdata", package = "IDConverter"))
```

### TCGA

```{r}
x <- convert_tcga("TCGA-02-0001-10")
x
```

### PCAWG

```{r}
x <- convert_pcawg("SP1677")
x
```

### ICGC

```{r}
x <- convert_icgc("SP29019")
x
```

### Genes

```{r}
convert_hm_genes(c("TP53", "KRAS", "EGFR", "MYC"), type = "symbol")

# Or use data from annotables
ls_annotables()
grch37 = load_data("grch37")
head(grch37)
convert_custom(c("TP53", "KRAS", "EGFR", "MYC"),
               from = "symbol", to = "entrez", dt = grch37)
```

### Human-Mouse Orthologs

```{r eval=FALSE}
# Human → Mouse
convert_hm_orthologs(c("TP53", "KRAS", "EGFR", "MYC"))
# => Trp53, Kras, Egfr, Myc

# Mouse → Human
convert_hm_orthologs(c("Trp53", "Kras"), from_species = "mouse", to_species = "human")
# => TP53, KRAS

# Ensembl ID mapping
convert_hm_orthologs("ENSG00000141510", from_type = "ensembl", to_type = "ensembl")
# => ENSMUSG00000059552
```

### Gene Symbol Aliases

```{r eval=FALSE}
# Build annotation table with synonym support
ann <- build_annotables("grch38", include_synonyms = TRUE, tx2gene = FALSE)

# Resolve aliases (e.g., old names → official symbols)
resolve_gene_aliases(c("TP53", "T245", "MLL"), ann[[1]])
# T245 → TSPAN6, MLL → KMT2A (if in synonym data)
```

### GDC Manifest Pairing

```{r eval=FALSE}
# Pair tumor-normal samples from a GDC manifest file
info <- pair_gdc_samples("gdc_manifest.txt")
head(info)
```

## Citation

***Wang S, Li H, Song M, Tao Z, Wu T, He Z, et al. (2021) Copy
number signature analysis tool and its application in prostate
cancer reveals distinct mutational processes and clinical outcomes.
PLoS Genet 17(5): e1009557.***
<https://doi.org/10.1371/journal.pgen.1009557>

## Similar package

- [AnnoProbe](https://github.com/jmzeng1314/AnnoProbe/) ([Gitee mirror](https://gitee.com/jmzeng/annoprobe)) is an R package for transforming Chips probes to different Gene IDs.

## LICENSE

MIT\@2020, Shixiang Wang
